### 2022-04-16 proposal/proposal.tex
Add proposal and papers.

\documentclass{amsart}

\usepackage{hyperref}

\title{COSC-373: Final Project Proposal}
\author{Lee Jiaen}
\author{Hyery Yoo}
\author{Alexander Lee}

\begin{document}

\maketitle

Recently, there has been a lot of hype about Bitcoin and other related
cryptocurrencies. We have therefore decided to investigate Bitcoin and its
blockchain technology for the final project. As a resource, we will read the
original Bitcoin paper by Satoshi Nakamoto
(\url{https://bitcoin.org/bitcoin.pdf}).

\end{document}

### 2022-04-29 notes/notes.tex
Start presentation notes.

\documentclass{article}

\title{Bitcoin Presentation Notes}
\author{}
\date{}

\begin{document}

\maketitle

\section{Introduction}

\begin{itemize}
  \item Commerce on the Internet relies almost exclusively on financial
    institutions serving as trusted third parties to process electronic
    payments.
  \item Problem: TODO
\end{itemize}

\section{Digital Signatures}

\begin{itemize}
  \item Scenario: Alice wants to send a message to Bob over a network. How can
    Bob verify that the message he received was from Alice? Alice needs to sign
    her message.
  \item Desired property of signature: cannot be forged on a different message.
  \item Both Alice and Bob each generate a public/private key pair for
    themselves. Each key is some string of bits. Private keys are kept secret.
  \item Producing a signature: $\mathsf{sign(message, privateKey)} =
    \mathsf{signature}$.
  \item Verifying a signature: $\mathsf{verify(message, signature, publicKey) =
    true / false}$.
  \item Only the owner of the private key can produce the signature.
  \item No one can copy the signature and forge it on another message.
  \item Signature is a 256 bit value. Hard to find a valid signature if you
    don't know the secret key. There is no strategy better than guessing and
    checking if random signatures are valid using the public key. There are
    $2^{256}$ signatures to check; this is a very large number.
\end{itemize}

\section{Transactions}

\begin{itemize}
  \item Electronic coin: a chain of digital signatures.
  \item Alice transfers a coin to Bob:
    \begin{enumerate}
      \item Alice computes the hash of previous transaction and Bob's public
        key.
      \item Alice signs the hash using her private key.
      \item Alice adds her signature to the end of the coin.
      \item Bob verifies that Alice transferred a coin to him using Alice's
        public key.
    \end{enumerate}
  \item Problem: how to verify that Alice did not double-spend the coin?
\end{itemize}

\end{document}

### 2022-04-30 notes/notes.tex
Continue working on notes.

@@
 
-\section{Introduction}
+\section{Motivation}
 
 \begin{itemize}
-  \item Commerce on the Internet relies almost exclusively on financial
-    institutions serving as trusted third parties to process electronic
-    payments.
-  \item Problem: TODO
+  \item Situation: commerce on the Internet relies almost exclusively on
+    financial institutions serving as trusted third parties to process
+    electronic payments.
+  \item Problem: such third parties may not be trustworthy.
+  \item Goal: develop a decentralized version of electronic cash that would
+    allow online payments to be sent directly from one party to another without
+    going through a financial institution.
 \end{itemize}
@@
     \begin{enumerate}
-      \item Alice computes the hash of previous transaction and Bob's public
-        key.
+      \item Alice computes the hash of the previous transaction and Bob's public
+        key. % TODO: maybe describe transaction
       \item Alice signs the hash using her private key.
       \item Alice adds her signature to the end of the coin.
-      \item Bob verifies that Alice transferred a coin to him using Alice's
-        public key.
     \end{enumerate}
-  \item Problem: how to verify that Alice did not double-spend the coin?
+  \item Bob can verify that Alice transferred a coin to him using Alice's public
+    key.
+  \item Problem: how to verify that Alice did not double-spend the coin? For
+    example, Alice could have transferred a coin to Charlie and then transferred
+    the same coin to Bob. For our purposes, the earliest transaction is the one
+    that counts, so we should ignore later attempts to double-spend. As of now,
+    there is no mechanism preventing Alice from double-spending.
+  \item To accomplish this without a trusted party, transactions must be
+    publicly announced, and we need a system for participants to agree on a
+    single history of the order in which they were received. Bob needs proof
+    that at the time of the transaction, the majority of nodes agreed it was the
+    first received.
+  \item Solution: a distributed timestamp server with proof-of-work.
+\end{itemize}
+
+\section{Timestamp Serve}
+
+\begin{itemize}
+  \item Take a hash of a block of transactions to be timestamped and widely
+    publishing the hash.
+  \item The timestamp proves that the data must have existed at the time,
+    obviously, in order to get the hash.
+  % TODO
+\end{itemize}
+
+\section{Cryptographic Hash Functions}
+
+\begin{itemize}
+  \item $\mathsf{SHA256(message) = hash}$.
+  \item Input: message.
+  \item Output: String of 256 bits, known as the hash of the message.
+  \item The output looks random even though it isn't (i.e., the function always
+    gives the same output for a given input).
+  \item Slightly changing the input results in the output to change completely
+    in an unpredictable way.
+  \item Infusible to compute in the reverse direction. Given only the output,
+    finding the input that hashes to the output is extremely difficult. There is
+    no better method than to guess and check. Need to go through $2^{256}$
+    guesses.
+\end{itemize}
+
+\section{Proof-of-Work}
+
+\begin{itemize}
+  \item % TODO
+\end{itemize}
+
+\section{Network}
+
+\begin{itemize}
+  \item The network executes as follows:
+    \begin{enumerate}
+      \item New transactions are broadcast to all nodes.
+      \item Each node collects new transactions into a block.
+      \item Each node works on finding a difficult proof-of-work for its block.
+      \item When a node finds a proof-of-work, it broadcasts the block to all
+        nodes.
+      \item Nodes accept the block only if all transactions in it are valid and
+        not already spent.
+      \item Nodes express their acceptance of the block by working on creating
+        the next block in the chain, using the hash of the accepted block as the
+        previous hash.
+    \end{enumerate}
 \end{itemize}

### 2022-05-01 notes/notes.tex
Combine stuff from notes.

@@
     going through a financial institution.
+  %TODO: add more? How could third parties not be trustworthy?
 \end{itemize}
 
-\section{Digital Signatures}
+\section{Detour: Digital Signatures}
 
@@
   \item Desired property of signature: cannot be forged on a different message.
-  \item Both Alice and Bob each generate a public/private key pair for
+  \item Both Alice and Bob each generate a private/public key pair for
     themselves. Each key is some string of bits. Private keys are kept secret.
@@
 
+\section{Bitcoin}
+
+\begin{itemize}
+  \item Bitcoin network: a randomly connected overlay network of a few thousand
+    nodes, controlled by a variety of owners. All nodes perform the same
+    operations, i.e., it is homogeneous network and without central control.
+  \item Users generate private/public key pairs.
+  \item Address: an identifier derived from a user's public key to receive funds
+    in Bitcoin.
+  \item The Bitcoin network collaboratively tracks the balance in bitcoins of
+    each address.
+\end{itemize}
+
 \section{Transactions}
@@
 \begin{itemize}
-  \item Electronic coin: a chain of digital signatures.
-  \item Alice transfers a coin to Bob:
-    \begin{enumerate}
-      \item Alice computes the hash of the previous transaction and Bob's public
-        key. % TODO: maybe describe transaction
-      \item Alice signs the hash using her private key.
-      \item Alice adds her signature to the end of the coin.
-    \end{enumerate}
-  \item Bob can verify that Alice transferred a coin to him using Alice's public
-    key.
-  \item Problem: how to verify that Alice did not double-spend the coin? For
-    example, Alice could have transferred a coin to Charlie and then transferred
-    the same coin to Bob. For our purposes, the earliest transaction is the one
-    that counts, so we should ignore later attempts to double-spend. As of now,
-    there is no mechanism preventing Alice from double-spending.
-  \item To accomplish this without a trusted party, transactions must be
-    publicly announced, and we need a system for participants to agree on a
-    single history of the order in which they were received. Bob needs proof
-    that at the time of the transaction, the majority of nodes agreed it was the
-    first received.
-  \item Solution: a distributed timestamp server with proof-of-work.
+  \item Output: a tuple consisting of an amount in bitcoins and a spending
+    condition requiring a valid signature associated with the private key of an
+    address.
+    \begin{itemize}
+      \item Outputs exists in two states: unspent and spent.
+      \item Any output can be spent at most once.
+      \item The address balance is the sum of bitcoin amounts in unspent outputs
+        that are associated with the address.
+      \item The set of unspent transaction outputs (UTXO) and some additional
+        global parameters is the shared state of Bitcoin. Every node in the
+        Bitcon network holds a complete replica of that state. Local replicas
+        may temporarily diverge but consistency is eventually re-established.
+    \end{itemize}
+  \item Input: a tuple consisting of a reference to a previously created output
+    and a signature to the spending condition, proving that the transaction
+    creator has the permission to spend the referenced output.
+  \item Transaction: a data structure that describes the transfer of bitcoins
+    from spenders to recipients. The transaction consists of a number of inputs
+    and new outputs. The inputs result in the referenced outputs spent (removed
+    from the UTXO), and the new outputs being added to the UTXO.\@
+  \item A transaction from Alice to Bob:
+    \begin{itemize}
+      \item Sender: Alice.
+      \item Recipients: Bob.
+      \item Inputs: one or more tuples where each tuple consists of a reference
+        to one of Alice's previously created unspent outputs and Alice's
+        signature using her private key.
+      \item Output: a tuple consisting of the amount in bitcoins for Bob and a
+        spending condition requiring a valid signature associated with the
+        private key of Bob's address.
+    \end{itemize}
+  \item Transactions are broadcast in the Bitcoin network and processed by every
+    node that receives them.
+  \item The outputs of a transaction may assign less than the sum of inputs, in
+    which case the difference is called the transaction's fee. The fee is used
+    to incentivize other participants in the system.
+  \item Transactions are in one of two states: unconfirmed or confirmed. In
+    coming transactions from the broad cast are unconfirmed and added to a pool
+    of transactions called the memory pool.
 \end{itemize}
 
-\section{Timestamp Serve}
+\section{Doublespends}
 
 \begin{itemize}
-  \item Take a hash of a block of transactions to be timestamped and widely
-    publishing the hash.
-  \item The timestamp proves that the data must have existed at the time,
-    obviously, in order to get the hash.
-  % TODO
+  \item Problem: nothing prevents nodes to locally accept different transactions
+    that spend the same output. For example, the transaction from Alice to Bob
+    could have spent the same output as the transaction from Alice to Charlie.
+  \item Doublespend: a situation in which multiple transactions attempt to spend
+    the same output. Only one transaction can be valid since outputs can only be
+    spent once. When nodes accept different transactions in a doublespend, the
+    shared state becomes inconsistent.
+  \item Doublespends can result in an inconsistent state since the validity of
+    transactions depends on the order in which they arrive. If two conflicting
+    transactions are seen by a node, the node considers the first to be valid.
+    The second transaction is invalid since it tries to spend an output that is
+    already present. The order in which transactions are seen may not be the
+    same for all node, hence the inconsistent state.
+  \item If doublespends are not resolved, the shared state diverges. Therefore a
+    conflict resolution mechanism is needed to decide which of the conflicting
+    transactions is to be confirmed (accepted by everybody), to achieve eventual
+    consistency.
 \end{itemize}
 
-\section{Cryptographic Hash Functions}
+\section{Detour: Cryptographic Hash Functions}
 
@@
     in an unpredictable way.
-  \item Infusible to compute in the reverse direction. Given only the output,
+  \item Infeasible to compute in the reverse direction. Given only the output,
     finding the input that hashes to the output is extremely difficult. There is
@@
     guesses.
+  \item However, given the input message, the hash is fast to compute.
 \end{itemize}
 
-\section{Proof-of-Work}
+\section{Proof-of-Work and Blocks}
 
 \begin{itemize}
-  \item % TODO
+  \item Proof-of-Work (PoW): a mechanism that allows a party to prove to another
+    party that a certain amount of computational resource has been utilized for
+    a period of time.
+  \item Block: a data structure used to communicate incremental changes to the
+    local state of a node. A block consists of a list of transactions, a
+    reference to a previous block and a nonce. A block lists some transactions
+    the block creator (``miner'') has accepted to its memory-pool since the
+    previous block. A node finds and broadcasts a block when it finds a valid
+    nonce.
+  \item A valid nonce is a number such that the SHA256 hash of the block begins
+    with a required number of zero bits.
+  \item A node finds a valid nonce for a block by incrementing the nonce from
+    zero and checking if the SHA256 hash begins with the required number of zero
+    bits.
+  \item Once a node finds a valid nonce for a block, the block cannot be changed
+    without redoing the work. As later blocks are chained after it, the work to
+    change the block would include redoing all the blocks after it.
+  \item The PoW also solves the problem of determining representation in
+    majority decision making. The majority decision is represented by the
+    longest chain, which has the greatest PoW effort invested in it.
 \end{itemize}
@@
     \end{enumerate}
+  \item Nodes always consider the longest chain to be the correct one and will
+    keep working on extending it.
+  \item If two nodes broadcast different versions of the next block
+    simultaneously, some nodes may receive one or the other first. In that case,
+    they work on the first one they received but save the other branch in case
+    it becomes longer; the nodes that were working on the other branch will then
+    switch to the longer one.
+  \item New transaction broadcasts do not necessarily need to reach all nodes.
+    As long as they reach many nodes, they will get into a block before long.
+  \item Block broadcasts are also tolerant of dropped messages. If a node does
+    not receive a block, it will request it when it receives the next block.
 \end{itemize}

### 2022-05-10 paper/paper.tex
Add paper skeleton.

\documentclass{article}

\usepackage{hyperref}

\title{COSC-373: Bitcoin}
\author{Lee Jiaen, Hyery Yoo, and Alexander Lee}

\begin{document}

\maketitle

\section{Introduction}

\section{Preliminaries}

\subsection{Digital Signatures}

\subsection{Cryptographic Hash Functions}

\section{Bitcoin}

\section{Transactions}

\section{Doublespends}

\section{Proof-of-Work}

\section{Network}

\section{Conclusion}

\end{document}

### 2022-05-10 paper/paper.tex
First pass at Digital Signatures.

@@
 
+To understand Bitcoin, we first need to develop an understanding of the
+fundamental building-blocks behind it: digital signatures and cryptographic hash
+functions.
+
 \subsection{Digital Signatures}
+
+Suppose for example that Alice wants to send a message to Bob over a network.
+How can Bob ensure that the message he received was from Alice and not from a
+malicious actor like Eve? What Alice can do is append a \emph{digital signature}
+to her message and send the message and signature to Bob, which allows Bob to
+verify that the message he received was indeed from Alice.
+
+The main idea behind digital signatures is as follows. Both Alice and Bob each
+generate a private/public key pair for themselves, where each key is a string of
+bits and private keys are kept secret (i.e., only Alice knows her private key
+and only Bob knows his private key). Producing a signature involves a function
+$\mathsf{sign(msg, priKey)}$, which takes Alice's message $\mathsf{msg}$ and
+private key $\mathsf{priKey}$ as inputs and outputs Alice's signature
+$\mathsf{sig}$. Similarly, verifying the authenticity of the received message
+involves a function $\mathsf{verify(msg, sig, pubKey)}$, which takes the
+received message $\mathsf{msg}$, the signature $\mathsf{sig}$, and Alice's
+public key $\mathsf{pubKey}$ as inputs and outputs \emph{true} if the signature
+was produced using Alice's message and private key, and \emph{false} otherwise.
+We will not discuss how the $\mathsf{sign()}$ and $\mathsf{verify()}$ functions
+work in depth since such details are not the focus of this work.
+
+Digital signatures have two required properties. First, the authenticity of
+Alice's signature generated from her message and private key can be verified
+easily using her corresponding public key. Secondly, it should be
+computationally infeasible for someone like Eve to generate a valid signature
+for Alice without knowing Alice's private key. That is, Eve has no better
+strategy than guessing and checking if random signatures are valid using Alice's
+message and public key. Assuming a 256 bit signature, the aforementioned brute
+force strategy would require Eve to check $2^{256}$ signatures in the worst
+case.
 

### 2022-05-11 paper/paper.tex
Add overview of Bitcoin.

@@
 
-\title{COSC-373: Bitcoin}
+\title{COSC-373 Final Report: Bitcoin}
 \author{Lee Jiaen, Hyery Yoo, and Alexander Lee}
@@
 
+The \emph{Bitcoin network} is a randomly connected overlay network of a few
+thousand nodes, where nodes are controlled by various owners and perform the
+same operation. In other words, the network is a homogeneous network without
+central control. Each user of Bitcoin generates a private/public key pair. A
+user is identified by their \emph{address}, which is derived from their public
+key and used to receive funds in Bitcoin. All nodes in the Bitcoin network work
+together to track each address' balance in bitcoins.
+
 \section{Transactions}

### 2022-05-12 paper/paper.tex
Add some emph.

@@
 The main idea behind digital signatures is as follows. Both Alice and Bob each
-generate a private/public key pair for themselves, where each key is a string of
-bits and private keys are kept secret (i.e., only Alice knows her private key
-and only Bob knows his private key). Producing a signature involves a function
-$\mathsf{sign(msg, priKey)}$, which takes Alice's message $\mathsf{msg}$ and
-private key $\mathsf{priKey}$ as inputs and outputs Alice's signature
-$\mathsf{sig}$. Similarly, verifying the authenticity of the received message
-involves a function $\mathsf{verify(msg, sig, pubKey)}$, which takes the
+generate a \emph{private key} and \emph{public key} pair for themselves, where
+each key is a string of bits and private keys are kept secret (i.e., only Alice
+knows her private key and only Bob knows his private key). Producing a signature
+involves a function $\mathsf{sign(msg, priKey)}$, which takes Alice's message
+$\mathsf{msg}$ and private key $\mathsf{priKey}$ as inputs and outputs Alice's
+signature $\mathsf{sig}$. Similarly, verifying the authenticity of the received
+message involves a function $\mathsf{verify(msg, sig, pubKey)}$, which takes the
 received message $\mathsf{msg}$, the signature $\mathsf{sig}$, and Alice's
@@
 Digital signatures have two required properties. First, the authenticity of
-Alice's signature generated from her message and private key can be verified
-easily using her corresponding public key. Secondly, it should be
-computationally infeasible for someone like Eve to generate a valid signature
-for Alice without knowing Alice's private key. That is, Eve has no better
-strategy than guessing and checking if random signatures are valid using Alice's
-message and public key. Assuming a 256 bit signature, the aforementioned brute
-force strategy would require Eve to check $2^{256}$ signatures in the worst
-case.
+Alice's signature generated from her message and private key can be
+\emph{verified easily} using her corresponding public key. Secondly, it should
+be \emph{computationally infeasible} for someone like Eve to generate a valid
+signature for Alice without knowing Alice's private key. That is, Eve has no
+better strategy than guessing and checking if random signatures are valid using
+Alice's message and public key. Assuming a 256 bit signature, the aforementioned
+brute force strategy would require Eve to check $2^{256}$ signatures in the
+worst case.
 

### 2022-05-12 paper/paper.tex
Start cryptographic hash functions.

--- 2022-05-11 paper/paper.tex
@@
 The main idea behind digital signatures is as follows. Both Alice and Bob each
-generate a private/public key pair for themselves, where each key is a string of
-bits and private keys are kept secret (i.e., only Alice knows her private key
-and only Bob knows his private key). Producing a signature involves a function
-$\mathsf{sign(msg, priKey)}$, which takes Alice's message $\mathsf{msg}$ and
-private key $\mathsf{priKey}$ as inputs and outputs Alice's signature
-$\mathsf{sig}$. Similarly, verifying the authenticity of the received message
-involves a function $\mathsf{verify(msg, sig, pubKey)}$, which takes the
+generate a \emph{private key} and \emph{public key} pair for themselves, where
+each key is a array of bits and private keys are kept secret (i.e., only Alice
+knows her private key and only Bob knows his private key). Producing a signature
+involves a function $\mathsf{sign(msg, priKey)}$, which takes Alice's message
+$\mathsf{msg}$ and private key $\mathsf{priKey}$ as inputs and outputs Alice's
+signature $\mathsf{sig}$. Similarly, verifying the authenticity of the received
+message involves a function $\mathsf{verify(msg, sig, pubKey)}$, which takes the
 received message $\mathsf{msg}$, the signature $\mathsf{sig}$, and Alice's
@@
 \subsection{Cryptographic Hash Functions}
+
+A \emph{cryptographic hash function} is a hash function that takes as input data
+(i.e., a \emph{message}) of an arbitrary size and outputs a bit array (i.e.,
+\emph{hash}) of a fixed size. In the case of the SHA256 cryptographic hash
+function, the hash has a length of 256 bits.
+
+Cryptographic hash functions ideally should have the following properties.
+First, computing the hash of a given message should be quick. Second, generating
+the message from only the hash should be computationally infeasible. Third,
+finding two different messages that map to the same hash under the function
+should also be infeasible. Fourth, slightly changing the message should change
+the hash so much such that the old and new hashes seem uncorrelated.
 

### 2022-05-12 paper/paper.tex
Fix wording.

@@
 A \emph{cryptographic hash function} is a hash function that takes as input data
-(i.e., a \emph{message}) of an arbitrary size and outputs a bit array (i.e.,
-\emph{hash}) of a fixed size. In the case of the SHA256 cryptographic hash
-function, the hash has a length of 256 bits.
+of an arbitrary size, known as the \emph{message}, and outputs a bit array of a
+fixed size, known as the \emph{hash}. In the case of the SHA256 cryptographic
+hash function, the hash has a length of 256 bits.
 

### 2022-05-17 paper/paper.tex
Finish cryptographic hash functions.
Update paper.tex

@@
 \section{Introduction}
+Before Bitcoin, the only commerce system on the internet relied heavily on trusted third parties to process electronic payments.
+Suppose Alice, a buyer, wants to buy an item from Bob's e-commerce store.
+Alice would need to make an online payment that relies on a financial institution or some trusted third party to process the transaction.
+However, Alice may not trust such third parties.
+Alice wishes for a way where her online payment can be sent directly to Bob without going through a third party.
+While the previous system worked for most transactions, it had a few significant inefficiency problems.
+Trusted third-party financial intuitions cannot avoid mediating disputes, making it impossible to have completely non-reversible transactions.
+Therefore, the need for trust increases, and "a certain percentage of fraud is accepted as unavoidable" (Nakamoto 2008).
+One could avoid these issues with the use of physical currency.
+However, like Alice, many people would prefer a decentralized version of electronic cash that allows any two willing parties to transact directly without a trusted third party.
+This is the goal and motivation of Bitcoin.
+The following paper describes how Bitcoin, an electronic payment system, is based on cryptographic proof instead of trust, allowing for direct transactions without trusted third parties or financial institutions.
 
@@
 should also be infeasible. Fourth, slightly changing the message should change
-the hash so much such that the old and new hashes seem uncorrelated.
+the hash so much such that the old and new hashes seem uncorrelated. Similar to
+digital signatures, the only way to find a message that produces a given hash is
+to via a brute force approach of guessing and checking possible messages. Again,
+assuming a 256 bit hash, this strategy would require $2^{256}$ hashes to check
+in the worst case.
 
@@
 
-\section{Doublespends}
+Suppose we have a \emph{transaction}, where Alice is the sender and Bob is the recipient. What is a transaction, and how does it work?
+
+A \emph{transaction} is a data structure that describes the transfer of bitcoins from spenders to recipients.
+The transaction consists of several inputs and outputs.
+More specifically, the inputs are a tuple consisting of a reference to a previously created output and a signature to the spending condition, proving that the transaction creator has the permission to spend the referenced output.
+Meanwhile, the outputs are a tuple composed of an amount in bitcoins and a spending condition requiring a valid signature associated with the private key of an address.
+The outputs of a transaction may assign less than the sum of inputs, in which case the difference is called the transaction fee.
+The transaction fee is used to incentivize other participants in the system.
+
+An output has two states, unspent and spent, and can only be spent at most once.
+When we sum up the bitcoin amounts of unspent outputs associated with an address, we get the address balance, and the set of these unspent transaction outputs is the shared state of Bitcoin.
+
+Transactions can be in one of two states: unconfirmed or confirmed.
+Unconfirmed transactions are incoming transactions from the broadcast that are added to a pool of transactions called the memory pool.
+All transactions are broadcasted to the entire Bitcoin network, with each node deciding whether or not to accept the transaction.
+As a result, every node in the Bitcoin network holds a complete replica of the shared state.
+The local replicas of this shared state may temporarily diverge, but consistency is eventually reestablished.
+
+Thus, for the transaction between Alice to Bob, the inputs are Alice's previous unspent outputs and Alice's digital signature, which was created using her private key.
+Meanwhile, the outputs of this transaction are the amount in Bitcoins for Bob and the spending condition requiring a valid signature associated with the private key of Bob's address.
+The transaction is then broadcasted in the Bitcoin network and processed by every node that receives it.
+
+\section{Doublespending Problem}
+
+Suppose we now have another transaction from Alice to Charlie that attempts to spend the same output as the transaction from Alice to Bob.
+\emph{Doublespending} is a situation in which multiple transactions attempt to spend the same output.
+However, only one transaction can be valid because outputs can only be spent at most once.
+Doublespending occurs because the order in which transactions are seen may not be the same for all nodes, and the validity of transactions depends on the order in which the transactions arrive.
+If a node sees two conflicting transactions, it considers only the first transaction it sees valid and the second invalid.
+The second seen transaction is invalid since it tries to spend an already present output.
+As a result, different nodes in the network accept different transactions, making their shared states inconsistent.
+If the problem of doublespending is not resolved, the shared states of the Bitcoin network diverge.
+The problem is that there is nothing preventing nodes from locally accepting different transactions that spend the same output.
+Therefore, a conflict resolution mechanism is needed to decide which of the conflicting transactions (only one transaction) should be accepted by every node to achieve eventual consistency in the network.
 
@@
 
+\section{References}
+https://bitcoin.org/bitcoin.pdf 
 \end{document}

### 2022-05-19 paper/paper.tex
Fix formatting and add bibtex.
Remaining sections added.

@@
 
-\usepackage{hyperref}
+\usepackage{hyperref} % for reference links
+\usepackage{natbib} % for bibliography
 
@@
 \section{Introduction}
-Before Bitcoin, the only commerce system on the internet relied heavily on trusted third parties to process electronic payments.
-Suppose Alice, a buyer, wants to buy an item from Bob's e-commerce store.
-Alice would need to make an online payment that relies on a financial institution or some trusted third party to process the transaction.
-However, Alice may not trust such third parties.
-Alice wishes for a way where her online payment can be sent directly to Bob without going through a third party.
-While the previous system worked for most transactions, it had a few significant inefficiency problems.
-Trusted third-party financial intuitions cannot avoid mediating disputes, making it impossible to have completely non-reversible transactions.
-Therefore, the need for trust increases, and "a certain percentage of fraud is accepted as unavoidable" (Nakamoto 2008).
-One could avoid these issues with the use of physical currency.
-However, like Alice, many people would prefer a decentralized version of electronic cash that allows any two willing parties to transact directly without a trusted third party.
-This is the goal and motivation of Bitcoin.
-The following paper describes how Bitcoin, an electronic payment system, is based on cryptographic proof instead of trust, allowing for direct transactions without trusted third parties or financial institutions.
+Before Bitcoin, the only commerce system on the internet relied heavily on
+trusted third parties to process electronic payments. Suppose Alice, a buyer,
+wants to buy an item from Bob's e-commerce store. Alice would need to make an
+online payment that relies on a financial institution or some trusted third
+party to process the transaction. However, Alice may not trust such third
+parties. Alice wishes for a way where her online payment can be sent directly to
+Bob without going through a third party. While the previous system worked for
+most transactions, it had a few significant inefficiency problems. Trusted
+third-party financial intuitions cannot avoid mediating disputes, making it
+impossible to have completely non-reversible transactions. Therefore, the need
+for trust increases, and ``a certain percentage of fraud is accepted as
+unavoidable''~\citep{Nakamoto08}. One could avoid these issues with the use of
+physical currency. However, like Alice, many people would prefer a decentralized
+version of electronic cash that allows any two willing parties to transact
+directly without a trusted third party. This is the goal and motivation of
+Bitcoin. The following paper describes how Bitcoin, an electronic payment
+system, is based on cryptographic proof instead of trust, allowing for direct
+transactions without trusted third parties or financial institutions.
 
@@
 
-Suppose we have a \emph{transaction}, where Alice is the sender and Bob is the recipient. What is a transaction, and how does it work?
+Suppose we have a \emph{transaction}, where Alice is the sender and Bob is the
+recipient. What is a transaction, and how does it work?
 
-A \emph{transaction} is a data structure that describes the transfer of bitcoins from spenders to recipients.
-The transaction consists of several inputs and outputs.
-More specifically, the inputs are a tuple consisting of a reference to a previously created output and a signature to the spending condition, proving that the transaction creator has the permission to spend the referenced output.
-Meanwhile, the outputs are a tuple composed of an amount in bitcoins and a spending condition requiring a valid signature associated with the private key of an address.
-The outputs of a transaction may assign less than the sum of inputs, in which case the difference is called the transaction fee.
-The transaction fee is used to incentivize other participants in the system.
+A \emph{transaction} is a data structure that describes the transfer of bitcoins
+from spenders to recipients. The transaction consists of several inputs and
+outputs. More specifically, the inputs are a tuple consisting of a reference to
+a previously created output and a signature to the spending condition, proving
+that the transaction creator has the permission to spend the referenced output.
+Meanwhile, the outputs are a tuple composed of an amount in bitcoins and a
+spending condition requiring a valid signature associated with the private key
+of an address. The outputs of a transaction may assign less than the sum of
+inputs, in which case the difference is called the transaction fee. The
+transaction fee is used to incentivize other participants in the system.
 
 An output has two states, unspent and spent, and can only be spent at most once.
-When we sum up the bitcoin amounts of unspent outputs associated with an address, we get the address balance, and the set of these unspent transaction outputs is the shared state of Bitcoin.
+When we sum up the bitcoin amounts of unspent outputs associated with an
+address, we get the address balance, and the set of these unspent transaction
+outputs is the shared state of Bitcoin.
 
-Transactions can be in one of two states: unconfirmed or confirmed.
-Unconfirmed transactions are incoming transactions from the broadcast that are added to a pool of transactions called the memory pool.
-All transactions are broadcasted to the entire Bitcoin network, with each node deciding whether or not to accept the transaction.
-As a result, every node in the Bitcoin network holds a complete replica of the shared state.
-The local replicas of this shared state may temporarily diverge, but consistency is eventually reestablished.
+Transactions can be in one of two states: unconfirmed or confirmed. Unconfirmed
+transactions are incoming transactions from the broadcast that are added to a
+pool of transactions called the memory pool. All transactions are broadcasted to
+the entire Bitcoin network, with each node deciding whether or not to accept the
+transaction. As a result, every node in the Bitcoin network holds a complete
+replica of the shared state. The local replicas of this shared state may
+temporarily diverge, but consistency is eventually reestablished.
 
-Thus, for the transaction between Alice to Bob, the inputs are Alice's previous unspent outputs and Alice's digital signature, which was created using her private key.
-Meanwhile, the outputs of this transaction are the amount in Bitcoins for Bob and the spending condition requiring a valid signature associated with the private key of Bob's address.
-The transaction is then broadcasted in the Bitcoin network and processed by every node that receives it.
+Thus, for the transaction between Alice to Bob, the inputs are Alice's previous
+unspent outputs and Alice's digital signature, which was created using her
+private key. Meanwhile, the outputs of this transaction are the amount in
+Bitcoins for Bob and the spending condition requiring a valid signature
+associated with the private key of Bob's address. The transaction is then
+broadcasted in the Bitcoin network and processed by every node that receives it.
 
@@
 
-Suppose we now have another transaction from Alice to Charlie that attempts to spend the same output as the transaction from Alice to Bob.
-\emph{Doublespending} is a situation in which multiple transactions attempt to spend the same output.
-However, only one transaction can be valid because outputs can only be spent at most once.
-Doublespending occurs because the order in which transactions are seen may not be the same for all nodes, and the validity of transactions depends on the order in which the transactions arrive.
-If a node sees two conflicting transactions, it considers only the first transaction it sees valid and the second invalid.
-The second seen transaction is invalid since it tries to spend an already present output.
-As a result, different nodes in the network accept different transactions, making their shared states inconsistent.
-If the problem of doublespending is not resolved, the shared states of the Bitcoin network diverge.
-The problem is that there is nothing preventing nodes from locally accepting different transactions that spend the same output.
-Therefore, a conflict resolution mechanism is needed to decide which of the conflicting transactions (only one transaction) should be accepted by every node to achieve eventual consistency in the network.
+Suppose we now have another transaction from Alice to Charlie that attempts to
+spend the same output as the transaction from Alice to Bob.
+\emph{Doublespending} is a situation in which multiple transactions attempt to
+spend the same output. However, only one transaction can be valid because
+outputs can only be spent at most once. Doublespending occurs because the order
+in which transactions are seen may not be the same for all nodes, and the
+validity of transactions depends on the order in which the transactions arrive.
+If a node sees two conflicting transactions, it considers only the first
+transaction it sees valid and the second invalid. The second seen transaction is
+invalid since it tries to spend an already present output. As a result,
+different nodes in the network accept different transactions, making their
+shared states inconsistent. If the problem of doublespending is not resolved,
+the shared states of the Bitcoin network diverge. The problem is that there is
+nothing preventing nodes from locally accepting different transactions that
+spend the same output. Therefore, a conflict resolution mechanism is needed to
+decide which of the conflicting transactions (only one transaction) should be
+accepted by every node to achieve eventual consistency in the network.
 
 \section{Proof-of-Work}
+Bitcoin solves the double-spending problem using \emph{Proof-of-Work} (PoW).
+PoW is a mechanism in which one party can prove to another that it has spent a specific amount of computational effort by presenting a nonce $x$ (a bitstring) to the PoW function $F_d(c,x)$, where $d$ is the difficulty (a positive number) and $c$ is the challenge (a bitstring).
+Another party can confirm this PoW by quickly checking that $F_d(c,x) =$ true for the given $x$.
+In general terms, suppose that Alice and Bob both know a PoW function $F_d(c,x)$ with $d$ and $c$ given.
+Alice expends a certain amount of computation to find a $x$ such that $F_d(c,x) =$ true.
+Alice presents her solution $x$ to Bob, and Bob checks the solution by computing $F_d(c,x)$.
+The PoW function must be defined such that verifying a solution $x$ is fast.
+Through this process, Alice proves to Bob that she has expended the amount of computational work specified by difficulty $d$.
+
+What does PoW look like in the context of Bitcoin?
+First, we must understand blocks.
+A \emph{block} is a data structure that consists of a list of transactions, a reference to a previous block, and a nonce.
+Reference to previous block links blocks together such that they form a blockchain.
+A block can be added to a chain by a node (called \emph{mining}) when it collects transactions from the network and finds a valid nonce $x$ that satisfies the Bitcoin PoW function:\newline
+$F_d (c,x) = SHA256(SHA256(c|x)) < 2^{224}/d$\newline
+Simply put, a node must find a value $x$ that, when concatenated with a node-specific challenge $c$ and hashed with the hash function SHA256 twice, returns a hash that starts with a fixed number of zeroes.
+SHA256 is a cryptographic hash function, so a valid nonce can only be found by iterating over all possible values of $x$, requiring significant computational work to be put in.
+Suppose Alice is a node who puts in the computational effort to find a valid nonce $x$ and broadcasts the block containing $x$ to other nodes.
+Alice’s PoW can be verified by another node Bob who sees Alice’s block.
+Bob is given the values of $c$ and $x$, so he can quickly compute the hash and verify that the result starts with the required number of zeroes.
 
 \section{Network}
+The Bitcoin network operates by each node in the network executing the same set of protocols.
+All transactions are broadcasted to the network, and every node works on collecting new transactions from the network into a block and finding a valid nonce for the block.
+When a node finds the valid nonce, it broadcasts the block to the network, adding the block to the end of the chain in effect.
+When another node in the network learns about the new block, it checks that all transactions in the block are valid.
+If they are, it accepts the block and works an extending the chain by creating a new block with the hash of the accepted block.
+
+A goal of the Bitcoin network is for all its nodes to agree on a single state of the network.
+The nodes achieve this by accepting the longest chain they see as the true state of the network.
+If a node receives two conflicting versions of a block, it works on extending the first block it receives and saves the other block as a branch.
+If the node later learns that the branch has become longer than the chain it was working on, it switches to extending the branch.
+Because every block requires PoW in order to be added to the chain, the longest chain in the network represents the greatest amount of collective computational effort, i.e., majority decision.
+All nodes eventually receive and accept the longest chain, so the network reaches consistency.
+
+Let us now look at how the Bitcoin network solves the doublespending problem.
+Suppose Alice attempts to doublespend by creating two transactions containing the same output, and the two transactions are collected into two separate blocks, Block 1 and Block 2.
+If one of the blocks, Block 1, is broadcasted to the network and gets accepted by all nodes before Block 2 is broadcasted, Block 2 is rejected by all nodes because it contains an invalid transaction.
+Thus, Alice’s doublespending attempt fails.
+In a more subtle case, suppose that Block 1 and Block 2 are broadcasted to the network (almost) simultaneously.
+Some of the nodes in the network receive Block 1 first, so they accept Block 1 and reject Block 2.
+The opposite is true for other nodes in the network.
+This creates two conflicting chains in the network with some of the nodes working on extending each of the chains.
+Suppose the majority of nodes accept Block 1.
+Then the sum of the computational effort put into Block 1’s chain is greater than that for Block 2’s chain, and Block 1’s chain grows faster.
+Eventually, every node receives this longer chain, and Block 2’s chain becomes rejected by all nodes, leaving only one of Alice’s transactions in the network.
 
 \section{Conclusion}
+We discussed in this report the fundamental structure of Bitcoin and how it solves the doublesepnding problem.
+Other aspects of Bitcoin include incentivizing miners, reclaiming unnecessary disk space, and combining and splitting values in a transaction.
+You can read more about these topics in the original Bitcoin paper [Nakamoto, 2008].
+Further work aims to increase the rate of transactions  by establishing micropayment channels between two parties outside the blockchain network (https://tik-old.ee.ethz.ch/file/716b955c130e6c703fac336ea17b1670/duplex-micropayment-channels.pdf).
+Lastly, the introduction of a decentralized system for financial transactions by Bitcoin led to a number of other blockchain-based cryptocurrencies such as Ethereum.
+You can read about the structure of Ethereum from Ferretti and D’Angelo’s 2019 paper (https://arxiv.org/pdf/1908.11808.pdf).
 
-\section{References}
-https://bitcoin.org/bitcoin.pdf 
+\bibliographystyle{plainnat}
+\bibliography{paper.bib}
+
 \end{document}

### 2022-05-19 paper/paper.tex
Changed citations. Need to fix formatting.

@@
 Other aspects of Bitcoin include incentivizing miners, reclaiming unnecessary disk space, and combining and splitting values in a transaction.
-You can read more about these topics in the original Bitcoin paper [Nakamoto, 2008].
-Further work aims to increase the rate of transactions  by establishing micropayment channels between two parties outside the blockchain network (https://tik-old.ee.ethz.ch/file/716b955c130e6c703fac336ea17b1670/duplex-micropayment-channels.pdf).
+You can read more about these topics in the original Bitcoin paper ~\citep{Nakamoto08}.
+Further work aims to increase the rate of transactions  by establishing micropayment channels between two parties outside the blockchain network ~\citep{Decker15}.
 Lastly, the introduction of a decentralized system for financial transactions by Bitcoin led to a number of other blockchain-based cryptocurrencies such as Ethereum.
-You can read about the structure of Ethereum from Ferretti and D’Angelo’s 2019 paper (https://arxiv.org/pdf/1908.11808.pdf).
+You can read about the structure of Ethereum from Ferretti and D’Angelo’s paper ~\citep{Ferretti19}.
 

### 2022-05-19 paper/paper.tex
Fix formatting and citations.

@@
 
+\usepackage{amsmath}
 \usepackage{hyperref} % for reference links
@@
 \section{Introduction}
+
 Before Bitcoin, the only commerce system on the internet relied heavily on
@@
 \section{Proof-of-Work}
-Bitcoin solves the double-spending problem using \emph{Proof-of-Work} (PoW).
-PoW is a mechanism in which one party can prove to another that it has spent a specific amount of computational effort by presenting a nonce $x$ (a bitstring) to the PoW function $F_d(c,x)$, where $d$ is the difficulty (a positive number) and $c$ is the challenge (a bitstring).
-Another party can confirm this PoW by quickly checking that $F_d(c,x) =$ true for the given $x$.
-In general terms, suppose that Alice and Bob both know a PoW function $F_d(c,x)$ with $d$ and $c$ given.
-Alice expends a certain amount of computation to find a $x$ such that $F_d(c,x) =$ true.
-Alice presents her solution $x$ to Bob, and Bob checks the solution by computing $F_d(c,x)$.
-The PoW function must be defined such that verifying a solution $x$ is fast.
-Through this process, Alice proves to Bob that she has expended the amount of computational work specified by difficulty $d$.
 
-What does PoW look like in the context of Bitcoin?
-First, we must understand blocks.
-A \emph{block} is a data structure that consists of a list of transactions, a reference to a previous block, and a nonce.
-Reference to previous block links blocks together such that they form a blockchain.
-A block can be added to a chain by a node (called \emph{mining}) when it collects transactions from the network and finds a valid nonce $x$ that satisfies the Bitcoin PoW function:\newline
-$F_d (c,x) = SHA256(SHA256(c|x)) < 2^{224}/d$\newline
-Simply put, a node must find a value $x$ that, when concatenated with a node-specific challenge $c$ and hashed with the hash function SHA256 twice, returns a hash that starts with a fixed number of zeroes.
-SHA256 is a cryptographic hash function, so a valid nonce can only be found by iterating over all possible values of $x$, requiring significant computational work to be put in.
-Suppose Alice is a node who puts in the computational effort to find a valid nonce $x$ and broadcasts the block containing $x$ to other nodes.
-Alice’s PoW can be verified by another node Bob who sees Alice’s block.
-Bob is given the values of $c$ and $x$, so he can quickly compute the hash and verify that the result starts with the required number of zeroes.
+Bitcoin solves the double-spending problem using \emph{Proof-of-Work} (PoW). PoW
+is a mechanism in which one party can prove to another that it has spent a
+specific amount of computational effort by presenting a nonce $x$ (a bitstring)
+to the PoW function $F_d(c,x)$, where $d$ is the difficulty (a positive number)
+and $c$ is the challenge (a bitstring). Another party can confirm this PoW by
+quickly checking that $F_d(c,x) =$ true for the given $x$. In general terms,
+suppose that Alice and Bob both know a PoW function $F_d(c,x)$ with $d$ and $c$
+given. Alice expends a certain amount of computation to find a $x$ such that
+$F_d(c,x) =$ true. Alice presents her solution $x$ to Bob, and Bob checks the
+solution by computing $F_d(c,x)$. The PoW function must be defined such that
+verifying a solution $x$ is fast. Through this process, Alice proves to Bob that
+she has expended the amount of computational work specified by difficulty $d$.
+
+What does PoW look like in the context of Bitcoin? First, we must understand
+blocks. A \emph{block} is a data structure that consists of a list of
+transactions, a reference to a previous block, and a nonce. Reference to
+previous block links blocks together such that they form a blockchain. A block
+can be added to a chain by a node (called \emph{mining}) when it collects
+transactions from the network and finds a valid nonce $x$ that satisfies the
+Bitcoin PoW function:
+\[
+  F_d (c,x) = \text{SHA256}(\text{SHA256}(c|x)) < \frac{2^{224}}{d}.
+\]
+Simply put, a node must find a value $x$ that, when concatenated with a
+node-specific challenge $c$ and hashed with the hash function SHA256 twice,
+returns a hash that starts with a fixed number of zeroes. SHA256 is a
+cryptographic hash function, so a valid nonce can only be found by iterating
+over all possible values of $x$, requiring significant computational work to be
+put in. Suppose Alice is a node who puts in the computational effort to find a
+valid nonce $x$ and broadcasts the block containing $x$ to other nodes. Alice’s
+PoW can be verified by another node Bob who sees Alice’s block. Bob is given the
+values of $c$ and $x$, so he can quickly compute the hash and verify that the
+result starts with the required number of zeroes.
 
 \section{Network}
-The Bitcoin network operates by each node in the network executing the same set of protocols.
-All transactions are broadcasted to the network, and every node works on collecting new transactions from the network into a block and finding a valid nonce for the block.
-When a node finds the valid nonce, it broadcasts the block to the network, adding the block to the end of the chain in effect.
-When another node in the network learns about the new block, it checks that all transactions in the block are valid.
-If they are, it accepts the block and works an extending the chain by creating a new block with the hash of the accepted block.
 
-A goal of the Bitcoin network is for all its nodes to agree on a single state of the network.
-The nodes achieve this by accepting the longest chain they see as the true state of the network.
-If a node receives two conflicting versions of a block, it works on extending the first block it receives and saves the other block as a branch.
-If the node later learns that the branch has become longer than the chain it was working on, it switches to extending the branch.
-Because every block requires PoW in order to be added to the chain, the longest chain in the network represents the greatest amount of collective computational effort, i.e., majority decision.
-All nodes eventually receive and accept the longest chain, so the network reaches consistency.
+The Bitcoin network operates by each node in the network executing the same set
+of protocols. All transactions are broadcasted to the network, and every node
+works on collecting new transactions from the network into a block and finding a
+valid nonce for the block. When a node finds the valid nonce, it broadcasts the
+block to the network, adding the block to the end of the chain in effect. When
+another node in the network learns about the new block, it checks that all
+transactions in the block are valid. If they are, it accepts the block and works
+an extending the chain by creating a new block with the hash of the accepted
+block.
+
+A goal of the Bitcoin network is for all its nodes to agree on a single state of
+the network. The nodes achieve this by accepting the longest chain they see as
+the true state of the network. If a node receives two conflicting versions of a
+block, it works on extending the first block it receives and saves the other
+block as a branch. If the node later learns that the branch has become longer
+than the chain it was working on, it switches to extending the branch. Because
+every block requires PoW in order to be added to the chain, the longest chain in
+the network represents the greatest amount of collective computational effort,
+i.e., majority decision. All nodes eventually receive and accept the longest
+chain, so the network reaches consistency.
 
 Let us now look at how the Bitcoin network solves the doublespending problem.
-Suppose Alice attempts to doublespend by creating two transactions containing the same output, and the two transactions are collected into two separate blocks, Block 1 and Block 2.
-If one of the blocks, Block 1, is broadcasted to the network and gets accepted by all nodes before Block 2 is broadcasted, Block 2 is rejected by all nodes because it contains an invalid transaction.
-Thus, Alice’s doublespending attempt fails.
-In a more subtle case, suppose that Block 1 and Block 2 are broadcasted to the network (almost) simultaneously.
-Some of the nodes in the network receive Block 1 first, so they accept Block 1 and reject Block 2.
-The opposite is true for other nodes in the network.
-This creates two conflicting chains in the network with some of the nodes working on extending each of the chains.
-Suppose the majority of nodes accept Block 1.
-Then the sum of the computational effort put into Block 1’s chain is greater than that for Block 2’s chain, and Block 1’s chain grows faster.
-Eventually, every node receives this longer chain, and Block 2’s chain becomes rejected by all nodes, leaving only one of Alice’s transactions in the network.
+Suppose Alice attempts to doublespend by creating two transactions containing
+the same output, and the two transactions are collected into two separate
+blocks, Block 1 and Block 2. If one of the blocks, Block 1, is broadcasted to
+the network and gets accepted by all nodes before Block 2 is broadcasted, Block
+2 is rejected by all nodes because it contains an invalid transaction. Thus,
+Alice’s doublespending attempt fails. In a more subtle case, suppose that Block
+1 and Block 2 are broadcasted to the network (almost) simultaneously. Some of
+the nodes in the network receive Block 1 first, so they accept Block 1 and
+reject Block 2. The opposite is true for other nodes in the network. This
+creates two conflicting chains in the network with some of the nodes working on
+extending each of the chains. Suppose the majority of nodes accept Block 1. Then
+the sum of the computational effort put into Block 1’s chain is greater than
+that for Block 2’s chain, and Block 1’s chain grows faster. Eventually, every
+node receives this longer chain, and Block 2’s chain becomes rejected by all
+nodes, leaving only one of Alice’s transactions in the network.
 
 \section{Conclusion}
-We discussed in this report the fundamental structure of Bitcoin and how it solves the doublesepnding problem.
-Other aspects of Bitcoin include incentivizing miners, reclaiming unnecessary disk space, and combining and splitting values in a transaction.
-You can read more about these topics in the original Bitcoin paper ~\citep{Nakamoto08}.
-Further work aims to increase the rate of transactions  by establishing micropayment channels between two parties outside the blockchain network ~\citep{Decker15}.
-Lastly, the introduction of a decentralized system for financial transactions by Bitcoin led to a number of other blockchain-based cryptocurrencies such as Ethereum.
-You can read about the structure of Ethereum from Ferretti and D’Angelo’s paper ~\citep{Ferretti19}.
+
+We discussed in this report the fundamental structure of Bitcoin and how it
+solves the doublesepnding problem. Other aspects of Bitcoin include
+incentivizing miners, reclaiming unnecessary disk space, and combining and
+splitting values in a transaction. You can read more about these topics in the
+original Bitcoin paper~\citep{Nakamoto08}. Further work aims to increase the
+rate of transactions  by establishing micropayment channels between two parties
+outside the blockchain network~\citep{Decker15}. Lastly, the introduction of a
+decentralized system for financial transactions by Bitcoin led to a number of
+other blockchain-based cryptocurrencies such as Ethereum~\citep{Buterin14}.
 

### 2022-05-20 paper/paper.tex
Make revisions.

@@
 \usepackage{amsmath}
-\usepackage{hyperref} % for reference links
-\usepackage{natbib} % for bibliography
+\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
+\SetArgSty{textnormal}
+\usepackage{hyperref}
+\usepackage{natbib}
+\usepackage{cleveref}
+\crefname{algorithm}{Alg.}{Algs.}
+\Crefname{algorithm}{Algorithm}{Algorithms}
 
@@
 
-Before Bitcoin, the only commerce system on the internet relied heavily on
-trusted third parties to process electronic payments. Suppose Alice, a buyer,
-wants to buy an item from Bob's e-commerce store. Alice would need to make an
-online payment that relies on a financial institution or some trusted third
+Before Bitcoin, e-commerce on the internet relied heavily on trusted third
+parties to process electronic payments. For example, suppose Alice, a buyer,
+wants to purchase an item from Bob's e-commerce store. Alice would need to make
+an online payment that relies on a financial institution or some trusted third
 party to process the transaction. However, Alice may not trust such third
-parties. Alice wishes for a way where her online payment can be sent directly to
-Bob without going through a third party. While the previous system worked for
-most transactions, it had a few significant inefficiency problems. Trusted
-third-party financial intuitions cannot avoid mediating disputes, making it
-impossible to have completely non-reversible transactions. Therefore, the need
-for trust increases, and ``a certain percentage of fraud is accepted as
-unavoidable''~\citep{Nakamoto08}. One could avoid these issues with the use of
-physical currency. However, like Alice, many people would prefer a decentralized
-version of electronic cash that allows any two willing parties to transact
-directly without a trusted third party. This is the goal and motivation of
-Bitcoin. The following paper describes how Bitcoin, an electronic payment
-system, is based on cryptographic proof instead of trust, allowing for direct
-transactions without trusted third parties or financial institutions.
+parties and wish for a way where her online payment can be sent directly to
+Bob without going through a third party. While the previously mentioned payment
+system works for most transactions, it has a few significant problems with
+efficiency. Trusted third-party financial intuitions cannot avoid mediating
+disputes, making it impossible to have completely non-reversible transactions.
+Therefore, the need for trust increases, and ``a certain percentage of fraud is
+accepted as unavoidable''~\citep{Nakamoto08}. One could avoid these issues with
+the use of physical currency. However, like Alice, many people would prefer a
+decentralized version of electronic cash that allows any two willing parties to
+transact directly without a trusted third party. This is the motivation behind
+Bitcoin. This report describes how Bitcoin, an electronic payment system, is
+based on cryptographic proof instead of trust, allowing for direct transactions
+without trusted third parties or financial institutions.
 
@@
 
-To understand Bitcoin, we first need to develop an understanding of the
-fundamental building-blocks behind it: digital signatures and cryptographic hash
+To understand Bitcoin, we first need to develop an understanding of its
+fundamental building-blocks: digital signatures and cryptographic hash
 functions.
@@
 
-Suppose for example that Alice wants to send a message to Bob over a network.
+Suppose for instance that Alice wants to send a message to Bob over a network.
 How can Bob ensure that the message he received was from Alice and not from a
 malicious actor like Eve? What Alice can do is append a \emph{digital signature}
-to her message and send the message and signature to Bob, which allows Bob to
-verify that the message he received was indeed from Alice.
+to her message and send the message and signature to Bob, allowing Bob to verify
+that the message he received was indeed from Alice.
 
@@
 generate a \emph{private key} and \emph{public key} pair for themselves, where
-each key is a array of bits and private keys are kept secret (i.e., only Alice
+each key is an array of bits and private keys are kept secret (i.e., only Alice
 knows her private key and only Bob knows his private key). Producing a signature
@@
 We will not discuss how the $\mathsf{sign()}$ and $\mathsf{verify()}$ functions
-work in depth since such details are not the focus of this work.
+work in depth since such details are not the focus of this report.
 
@@
 digital signatures, the only way to find a message that produces a given hash is
-to via a brute force approach of guessing and checking possible messages. Again,
+via a brute force approach of guessing and checking possible messages. Again,
 assuming a 256 bit hash, this strategy would require $2^{256}$ hashes to check
@@
 same operation. In other words, the network is a homogeneous network without
-central control. Each user of Bitcoin generates a private/public key pair. A
+central control. Each user of Bitcoin generates a private/public key pair, and a
 user is identified by their \emph{address}, which is derived from their public
@@
 
-Suppose we have a \emph{transaction}, where Alice is the sender and Bob is the
-recipient. What is a transaction, and how does it work?
+Suppose we have a transaction in bitcoins, where Alice is the sender and Bob is
+the recipient. What is a transaction in this case?
 
 A \emph{transaction} is a data structure that describes the transfer of bitcoins
-from spenders to recipients. The transaction consists of several inputs and
-outputs. More specifically, the inputs are a tuple consisting of a reference to
-a previously created output and a signature to the spending condition, proving
-that the transaction creator has the permission to spend the referenced output.
-Meanwhile, the outputs are a tuple composed of an amount in bitcoins and a
+from senders to recipients. The transaction consists of several inputs and
+outputs. Each \emph{output} is a tuple composed of an amount in bitcoins and a
 spending condition requiring a valid signature associated with the private key
-of an address. The outputs of a transaction may assign less than the sum of
-inputs, in which case the difference is called the transaction fee. The
-transaction fee is used to incentivize other participants in the system.
+of an address. Meanwhile, each \emph{input} is a tuple consisting of a reference
+to a previously created output and a signature to the spending condition,
+proving that the transaction creator has the permission to spend the referenced
+output. An output has two states, \emph{unspent} and \emph{spent}, and can only
+be spent at most once. The sum of the unspent outputs associated with an address
+is the \emph{address balance}, and the set of these unspent transaction outputs
+(UTXO) is the shared state of Bitcoin. That is, every node in the Bitcoin
+network holds a complete replica of the shared state. The local replicas of this
+shared state may temporarily diverge, but consistency is eventually
+re-established. For a given transaction, the inputs result in the referenced
+outputs spent (i.e., removed from the UTXO), and the new outputs are added to
+the UTXO.\@ More specifically, inputs reference the output that is being spent
+by a $(h, i)$ tuple, where $h$ is the hash of the transaction that created the
+output, and $i$ is the index of the output in that transaction.
 
-An output has two states, unspent and spent, and can only be spent at most once.
-When we sum up the bitcoin amounts of unspent outputs associated with an
-address, we get the address balance, and the set of these unspent transaction
-outputs is the shared state of Bitcoin.
+All transactions are broadcasted to the Bitcoin network and processed by every
+node that receives them. Transactions can be in one of two states:
+\emph{unconfirmed} or \emph{confirmed}. Incoming transactions are unconfirmed
+and added to a pool of transactions referred to the \emph{memory pool}. We will
+discuss how transactions are confirmed in \cref{sec:pow}.
 
-Transactions can be in one of two states: unconfirmed or confirmed. Unconfirmed
-transactions are incoming transactions from the broadcast that are added to a
-pool of transactions called the memory pool. All transactions are broadcasted to
-the entire Bitcoin network, with each node deciding whether or not to accept the
-transaction. As a result, every node in the Bitcoin network holds a complete
-replica of the shared state. The local replicas of this shared state may
-temporarily diverge, but consistency is eventually reestablished.
+Returning to our example, for the transaction between Alice to Bob, the inputs
+are Alice's previous unspent outputs and Alice's digital signature, which was
+created using her private key. Meanwhile, the outputs of this transaction are
+the amount in Bitcoins for Bob and the spending condition requiring a valid
+signature associated with the private key of Bob's address. The transaction is
+then broadcasted to the Bitcoin network and processed by every node that
+receives it.
 
-Thus, for the transaction between Alice to Bob, the inputs are Alice's previous
-unspent outputs and Alice's digital signature, which was created using her
-private key. Meanwhile, the outputs of this transaction are the amount in
-Bitcoins for Bob and the spending condition requiring a valid signature
-associated with the private key of Bob's address. The transaction is then
-broadcasted in the Bitcoin network and processed by every node that receives it.
+\section{The Doublespending Problem}
 
-\section{Doublespending Problem}
-
-Suppose we now have another transaction from Alice to Charlie that attempts to
-spend the same output as the transaction from Alice to Bob.
-\emph{Doublespending} is a situation in which multiple transactions attempt to
+Now suppose there is a transaction from Alice to Charlie that attempts to spend
+the same output as the transaction from Alice to Bob. Observe that, for now,
+nothing prevents nodes in the Bitcoin network to locally accept different
+transactions that spend the same output, i.e., doublespend. More formally,
+\emph{doublespending} is a situation in which multiple transactions attempt to
 spend the same output. However, only one transaction can be valid because
 outputs can only be spent at most once. Doublespending occurs because the order
-in which transactions are seen may not be the same for all nodes, and the
+in which transactions are processed may not be the same for all nodes, and the
 validity of transactions depends on the order in which the transactions arrive.
-If a node sees two conflicting transactions, it considers only the first
-transaction it sees valid and the second invalid. The second seen transaction is
-invalid since it tries to spend an already present output. As a result,
-different nodes in the network accept different transactions, making their
-shared states inconsistent. If the problem of doublespending is not resolved,
-the shared states of the Bitcoin network diverge. The problem is that there is
-nothing preventing nodes from locally accepting different transactions that
-spend the same output. Therefore, a conflict resolution mechanism is needed to
-decide which of the conflicting transactions (only one transaction) should be
-accepted by every node to achieve eventual consistency in the network.
+If a node sees two conflicting transactions (i.e., transactions that spend the
+same output), the node only considers the first transaction it sees valid and
+the second invalid. The second transaction seen is invalid since it tries to
+spend an already spent output. As a result, different nodes in the network may
+accept different transactions, making their shared states inconsistent. If the
+problem of doublespending is not resolved, the shared states of the Bitcoin
+network diverge. Therefore, a conflict resolution mechanism is needed to decide
+which of the conflicting transactions should be accepted by every node to
+achieve eventual consistency in the network.
 
-\section{Proof-of-Work}
+\section{Processing a Transaction}
 
-Bitcoin solves the double-spending problem using \emph{Proof-of-Work} (PoW). PoW
-is a mechanism in which one party can prove to another that it has spent a
-specific amount of computational effort by presenting a nonce $x$ (a bitstring)
-to the PoW function $F_d(c,x)$, where $d$ is the difficulty (a positive number)
-and $c$ is the challenge (a bitstring). Another party can confirm this PoW by
-quickly checking that $F_d(c,x) =$ true for the given $x$. In general terms,
-suppose that Alice and Bob both know a PoW function $F_d(c,x)$ with $d$ and $c$
-given. Alice expends a certain amount of computation to find a $x$ such that
-$F_d(c,x) =$ true. Alice presents her solution $x$ to Bob, and Bob checks the
-solution by computing $F_d(c,x)$. The PoW function must be defined such that
-verifying a solution $x$ is fast. Through this process, Alice proves to Bob that
-she has expended the amount of computational work specified by difficulty $d$.
+Before we continue with discussing the solution to the doublespending problem,
+we first summarize how a node processes an incoming transaction (see
+\cref{algo:proctrans}). First, the node receives a transaction $t$
+(\cref{algline:rectrans}). Then, for each input $(h, i)$ in transaction $t$
+(where $h$ is the hash of the transaction that created the referenced output,
+and $i$ is the index of the output in that transaction), if $(h, i)$ is not in
+the node's local set of unspent transaction outputs or the signature for the
+output is invalid (\cref{algline:outputorsig}), then transaction $t$ is dropped
+and the node halts. Next, if the sum of the values of inputs is less than the
+sum of values of the new outputs (\cref{algline:sums}), then transaction $t$ is
+dropped and the node halts since one cannot spend more than they have. Then, for
+each input $(h, i)$ in transaction $t$, $(h, i)$ is removed from the node's
+local set of unspent transaction outputs (\cref{algline:remove}). Finally,
+transaction $t$ is appended to the node's local history of transactions and $t$
+is broadcasted to its neighbors in the Bitcoin network.
 
-What does PoW look like in the context of Bitcoin? First, we must understand
-blocks. A \emph{block} is a data structure that consists of a list of
-transactions, a reference to a previous block, and a nonce. Reference to
-previous block links blocks together such that they form a blockchain. A block
-can be added to a chain by a node (called \emph{mining}) when it collects
-transactions from the network and finds a valid nonce $x$ that satisfies the
-Bitcoin PoW function:
+\begin{algorithm}
+  \caption{Process Transaction}\label{algo:proctrans}
+  \DontPrintSemicolon%
+
+  Receive transaction $t$\;\label{algline:rectrans}
+  \ForEach{input $(h, i)$ in $t$}{
+    \If{output $(h, i)$ not in local UTXO \textbf{or} signature
+      invalid}{\label{algline:outputorsig}
+      Drop $t$ and stop\;
+    }
+  }
+  \If{sum of values of inputs $<$ sum of values of new
+    outputs}{\label{algline:sums}
+    Drop $t$ and stop\;
+  }
+  \ForEach{input $(h, i)$ in $t$}{
+    Remove $(h, i)$ from local UTXO\;\label{algline:remove}
+  }
+  Append $t$ to local history\;
+  Forward $t$ to neighbors in the Bitcoin network\;
+\end{algorithm}
+
+\section{Proof-of-Work}\label{sec:pow}
+
+Bitcoin solves the doublespending problem using \emph{Proof-of-Work} (PoW), a
+mechanism in which one party can prove to another that it has spent a specific
+amount of computational effort by presenting a nonce $x$ (a bit-string) to the
+PoW function $F_d(c,x) \to \{true, false\}$, where $d$ is the difficulty (a
+positive number) and $c$ is the challenge (a bit-string). Another party can
+confirm this PoW by quickly checking that $F_d(c,x) = true$ for the given $x$.
+For instance, suppose that Alice and Bob both know a PoW function $F_d(c,x)$
+with $d$ and $c$ given. Alice expends a certain amount of computation to find a
+$x$ such that $F_d(c,x) = true$. Alice presents her solution $x$ to Bob, and Bob
+checks the solution by computing $F_d(c,x)$ and seeing that the functions output
+is $true$. The PoW function must be defined such that verifying a solution $x$
+is fast. Through this process, Alice proves to Bob that she has expended the
+amount of computational work specified by difficulty $d$.
+
+What does PoW look like in the context of Bitcoin? First, we must understand the
+notion of a block in Bitcoin. A \emph{block} is a data structure that consists
+of a list of transactions, a reference to a previous block, and a nonce. The
+list of transactions are the transactions the node has accepted to its memory
+pool since the previous block. Thus, a block encodes incremental changes to the
+local state of the node. A node finds a block by finding a valid nonce $x$ that
+satisfies the Bitcoin PoW function:
 \[
-  F_d (c,x) = \text{SHA256}(\text{SHA256}(c|x)) < \frac{2^{224}}{d}.
+  F_d(c,x) = \text{SHA256}(\text{SHA256}(c|x)) < \frac{2^{224}}{d}.
 \]
-Simply put, a node must find a value $x$ that, when concatenated with a
-node-specific challenge $c$ and hashed with the hash function SHA256 twice,
-returns a hash that starts with a fixed number of zeroes. SHA256 is a
-cryptographic hash function, so a valid nonce can only be found by iterating
-over all possible values of $x$, requiring significant computational work to be
-put in. Suppose Alice is a node who puts in the computational effort to find a
-valid nonce $x$ and broadcasts the block containing $x$ to other nodes. Alice’s
-PoW can be verified by another node Bob who sees Alice’s block. Bob is given the
-values of $c$ and $x$, so he can quickly compute the hash and verify that the
-result starts with the required number of zeroes.
+This process is called \emph{mining}. Simply put, a node must find a value $x$
+that, when concatenated with a node-specific challenge $c$ and hashed twice with
+the SHA256 hash function, returns a hash that starts with a fixed number of
+zeroes. SHA256 is a cryptographic hash function, so a valid nonce can only be
+found by iterating over all possible values of $x$, requiring significant
+computational work. Once a node finds a valid nonce for the block, the block is
+broadcasted to the node's neighbors. Transactions contained in a block are said
+to be \emph{confirmed} by that block. The process for finding a block is
+summarized in \cref{algo:findblock}.
 
-\section{Network}
+To make concept of PoW more concrete, suppose Alice is a node who puts in the
+computational effort to find a valid nonce $x$ and broadcasts the block
+containing $x$ to other nodes. Alice’s PoW can be verified by another node Bob
+who sees Alice’s block. Bob is given the values of $c$ and $x$, so he can
+quickly compute the hash and verify that the result starts with the required
+number of zeroes.
 
-The Bitcoin network operates by each node in the network executing the same set
-of protocols. All transactions are broadcasted to the network, and every node
-works on collecting new transactions from the network into a block and finding a
-valid nonce for the block. When a node finds the valid nonce, it broadcasts the
-block to the network, adding the block to the end of the chain in effect. When
-another node in the network learns about the new block, it checks that all
-transactions in the block are valid. If they are, it accepts the block and works
-an extending the chain by creating a new block with the hash of the accepted
-block.
+\begin{algorithm}
+  \caption{Find Block}\label{algo:findblock}
+  \DontPrintSemicolon%
 
-A goal of the Bitcoin network is for all its nodes to agree on a single state of
-the network. The nodes achieve this by accepting the longest chain they see as
-the true state of the network. If a node receives two conflicting versions of a
-block, it works on extending the first block it receives and saves the other
+  \SetKwRepeat{Do}{do}{while}
+  \KwIn{Nonce $x = 0$, challenge $c$, difficulty $d$, previous block $b_{t-1}$}
+
+  \While{$F_d(c,x) = false$}{
+    $x = x + 1$\;
+  }
+  Broadcast block $b_t = (memory\ pool, b_{t-1}, x)$\;
+\end{algorithm}
+
+\section{Receive Block}
+
+A primary goal of the Bitcoin network is for all its nodes to agree on a single
+state of the network. To do this, the blocks build a tree with their reference
+to a previous block, where the tree's root is known as the \emph{genesis block}.
+The longest path from the genesis block to the leaf of the tree is called the
+\emph{blockchain}. Eventual consistency is achieved by accepting the longest
+path in the tree (i.e., the blockchain) as the true state of the network.
+Branches are formed if a node receives two conflicting versions of a block. In
+this case, it works on extending the first block it receives and saves the other
 block as a branch. If the node later learns that the branch has become longer
@@
 i.e., majority decision. All nodes eventually receive and accept the longest
-chain, so the network reaches consistency.
+chain, so the network reaches consistency. More formally,
+\cref{algo:receiveblock} describes how a node updates its local state once it
+receive a block. We define the height $h_b$ as the path length from the genesis
+block to block $b$. Namely, we denote the node's current head as block $b_{max}$
+with height $h_{max}$. First, the node receive block $b$
+(\cref{algline:recblock}). Then, the node connects block $b$ in the tree as the
+child of its parent $p$ at height $h_b = h_p + 1$. If the height $h_b$ of block
+$b$ is larger that the height $h_{max}$ of the node's current head at block
+$b_{max}$ (\cref{algline:height}), then the head block and height are updated
+appropriately, the node computes the UTXO for the path leading to the new head
+block $b_{max}$, and the memory pool is cleaned up.
+
+\begin{algorithm}
+  \caption{Receive Block}\label{algo:receiveblock}
+  \DontPrintSemicolon%
+
+  \KwIn{Node's current head block $b_{max}$ at height $h_{max}$}
+
+  Receive block $b$\;\label{algline:recblock}
+  Connect block $b$ in the tree as child of its parent $p$ at height $h_b = h_p +
+  1$\;
+  \If{$h_b > h_{max}$}{\label{algline:height}
+    $h_{max} = h_b$\;
+    $b_{max} = b$\;
+    Compute UTXO for the path leading to $b_{max}$\;
+    Cleanup memory pool\;
+  }
+\end{algorithm}
+
+\section{Network}
+
+To summarize, the Bitcoin network operates by each node in the network executing
+the same set of protocols. All transactions are broadcasted to the network, and
+every node works on collecting new transactions from the network into a block
+and finding a valid nonce for the block. When a node finds the valid nonce, it
+broadcasts the block to the network, adding the block to the end of the chain in
+effect. When another node in the network learns about the new block, it checks
+that all transactions in the block are valid. If they are, it accepts the block
+and works on extending the chain by creating a new block with the hash of the
+accepted block.
+
+\section{Correctness}
 
@@
 
-We discussed in this report the fundamental structure of Bitcoin and how it
-solves the doublesepnding problem. Other aspects of Bitcoin include
-incentivizing miners, reclaiming unnecessary disk space, and combining and
-splitting values in a transaction. You can read more about these topics in the
-original Bitcoin paper~\citep{Nakamoto08}. Further work aims to increase the
-rate of transactions  by establishing micropayment channels between two parties
-outside the blockchain network~\citep{Decker15}. Lastly, the introduction of a
-decentralized system for financial transactions by Bitcoin led to a number of
-other blockchain-based cryptocurrencies such as Ethereum~\citep{Buterin14}.
+This report discusses the fundamental structure of Bitcoin and how it solves the
+doublespending problem. Other aspects of Bitcoin include incentivizing miners,
+reclaiming unnecessary disk space, and scripting. We refer the reader to learn
+more about these topics from the original Bitcoin white paper~\citep{Nakamoto08}
+and lecture notes from ETH Zurich by~\citeauthor{ETH15}. Further work aims to
+increase the rate of transactions  by establishing micropayment channels between
+two parties outside the blockchain network~\citep{Decker15}. Moreover, the
+introduction of a decentralized system for financial transactions by Bitcoin led
+to a number of other blockchain-based cryptocurrencies such as
+Ethereum~\citep{Buterin14}.
 
